Using spatially biased variables in ecosystem condition accounting with a GIS based workflow

### 2024-02-26 manuscript/HIAs.qmd

ggsave("../images/spread-na.tiff",
       plot = spread_na_map)
ggsave("../images/spread-na-small.tiff",
       plot = spread_na_map,
       dpi=150)
ggsave("../images/spread-na-small.jpg",
       plot = spread_na_map)
```

@@

combineALl <- rbind(
  spread_nf,
  spread_gr,
  spread_na
) |>
  as_tibble() |>
  drop_na() |>
  group_by(Municipality, indicator) |>
  rowwise() |>
  mutate(
    i_sample = list(sample(rnorm(10, mean, sd)))) |>
  unnest(cols=i_sample) |>
  summarise(mean = mean(i_sample),
            lower = quantile(i_sample, probs = c(0.025)),
            upper = quantile(i_sample, probs = c(0.975)))

# add area weights to the above and make forest plot

#combineALl
#tail(combineALl)
#combineALl$i_sample[2]
```

@@

![Schematic workflow followed in this paper. Sub-index A is calculated using spatially explicit aggregation (i.e .a simple overlay taking the mean value).](../images/workflow.jpg){#fig-workflow}

@@

>>>>

#ggsave("../images/spread-na.tiff",
#       plot = spread_na_map)
#ggsave("../images/spread-na-small.tiff",
#       plot = spread_na_map,
#       dpi=150)
#ggsave("../images/spread-na-small.jpg",
#       plot = spread_na_map)
```

@@

# Now I sample indicator values from the EDM and create a new distribution for 
# indicator values in the EEA (the municipalities). First I sample the individual
# distributions for each mire polygon with n defined by the polygon area.
# The distribution for the polygon areas is strongly right skewed, meaning
# some polygons will contribute much more to the EAA value than others.
# I will keep this design in-line with SEEA EA guidelines for area weighting, 
# but it's worth noting that this could be solved in other ways.
# 
combineAll <- rbind(
  spread_nf,
  spread_gr,
  spread_na
) |>
  as_tibble() |>
  drop_na() |>
  group_by(Municipality, indicator) |>
  rowwise() |>
  mutate(
    i_sample = list(sample(rnorm(n, mean, sd)))) |>
  select(-geometry) |>
  group_by(Municipality, indicator) |>
  reframe(i_sample = 
            sample(
              unlist(i_sample), 
              size = 1000,
              replace = TRUE)) |>
  mutate(i_sample = case_when(
    i_sample < 0 ~ 0,
    i_sample >1  ~ 1,
    .default = i_sample
  ))
```

```{r EEA-plot}
#| eval: true
#| cache: true
library(ggridges)

(ridgepPlot <- combineAll |>
  mutate(
    indicator = case_when(
      indicator == "i_ADSV" ~ "ADSV",
      indicator == "i_alien" ~ "Alien species",
      indicator == "i_ditch" ~ "Trenching"
    )) |>
  ggplot(
    aes(
      x = i_sample, 
      y = Municipality,
      fill = after_stat(x))
  ) +
  geom_density_ridges_gradient(
    bandwidth=.05,
    quantile_lines = TRUE, 
    quantiles = c(0.025, .5, 0.975)
  ) +
  #scale_fill_viridis_c(option = "D") +
  scale_fill_distiller(palette = "RdYlGn",
    direction = 1) +
  theme_bw(base_size = 15) +
  guides(fill = "none") +
  geom_vline(xintercept = .6,
    size=1.2,
    lty=2) +
  theme(axis.text.x = element_text(angle=45, hjust=1, vjust=1))+
  labs(y = "", x = "Indicator values")+
  xlim(c(0,1)) +
  facet_wrap(.~indicator)
)

#ggsave(plot = ridgepPlot,
#       "../images/ridgePlot.jpg",
#       width=10,
#       height=5)
```

```{r EEA-tbl}
#| eval: true
(EEA_tbl <- combineAll |>
  mutate(
    indicator = case_when(
      indicator == "i_ADSV" ~ "ADSV",
      indicator == "i_alien" ~ "Alien species",
      indicator == "i_ditch" ~ "Trenching"
    )) |>
  group_by(Municipality, indicator) |>
  summarise(mean = round(mean(i_sample), 2),
            median = round(median(i_sample), 2),
            percentile_025 = round(quantile(i_sample, probs = 0.025), 2),
            percentile_975 = round(quantile(i_sample, probs = 0.975), 2))
)

# saveRDS(EEA_tbl, "../output/EEA-table.RDS")
```

@@

```{r fig-workflow}
#| eval: true
#| include: true
#| out.width: '90%'
#| fig.cap:
#| - "Schematic workflow followed in this paper. Sub-index A is calculated using 
#| spatially explicit aggregation."

knitr::include_graphics("../images/workflow.jpg")
```

@@

```{r fig-EEA-ridgePlot}
#| eval: true
#| include: true
#| out.width: '80%'
#| fig.cap:
#| - "Distributions for three ecosystem condition indicators in the Norwegian 
#| Municiaplities. The color gradient reflects the same as the x-axis. 
#| The dotted vertical line represnet the threshold for what is considered 
#| reduced ecosystem condtion (< 0.6). "

knitr::include_graphics("../images/ridgePlot.jpg")
```
